---
title: "FCD + BrainSpan Controls"
output: html_document
date: "`r Sys.Date()`"
---


```{r setup, include=FALSE}
rm(list = ls())

homedir <- Sys.getenv("google_drive_i199200")

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(readxl)
library(xlsx)
library(icesTAF)
```

# Data sets 

```{r loading_metadata_Dataset1, include = F}
setwd(paste0(homedir, "Meu Drive/Doutorado/Deconvulucao/data/metadata_padronizada"))

# Dataset 1 is the discovery cohort
meta_disc <- read_excel("std_meta_FCD_AusFabioFiltered.xlsx")

meta_Aus <- meta_disc %>% filter(is.na(patient))
meta_Fabio <- meta_disc %>% filter(!is.na(patient))


```


## Kobow et al., 2019


For this data set, we have the following configuration of patients:

```{r, warning = F}
counts <- meta_Aus %>% count(diagnosis)
kable(counts, col.names = c("Diagnosis", "Patients"), align = "cc", format = "latex", booktabs = T) %>% kable_styling(latex_options = c("striped", "HOLD_position"))
```

There are `r counts[counts$diagnosis == "FCD IIa", 2]` FCD IIa patients and `r counts[counts$diagnosis == "FCD IIb", 2]` FCD IIb patients.


For this dataset, we have information on the disease, diagnosis, sex, age, onset, duration and lesion localization (lobe and (for the majority) lateralization). 


## Assis-Mendonça et al., 2023

For this data set, we have the following configuration of patients:

```{r, warning = F}
counts <- meta_Fabio %>% count(diagnosis)
kable(counts, col.names = c("Diagnosis", "Patients"), align = "cc", format = "latex", booktabs = T) %>% kable_styling(latex_options = c("striped", "HOLD_position"))
```

There are `r counts[counts$diagnosis == "FCD IIa", 2]` FCD IIa patients and `r counts[counts$diagnosis == "FCD IIb", 2]` FCD IIb patients.


Additionally, for these patients we have information on the disease, diagnosis, sex, age, onset, and duration. For a subset of these patients (5/8), we also have information on the lobe (all Frontal).


## Zimmer et al., 2021

```{r loading_metadata_Dataset2, include = F}
setwd(paste0(homedir, "Meu Drive/Doutorado/Deconvulucao/data/metadata_padronizada"))
meta_val <- read_excel("std_meta_Holanda.xlsx") %>% filter(diagnosis %in% c("FCD IIa", "FCD IIb"))
```


For this data set, we have the following configuration of patients:

```{r}
counts <- meta_val %>% count(diagnosis)
kable(counts, col.names = c("Diagnosis", "Patients"), align = "cc", format = "latex", booktabs = T) %>% kable_styling(latex_options = c("striped", "HOLD_position"))
```

There are `r counts[counts$diagnosis == "FCD IIa", 2]` FCD IIa patients and `r counts[counts$diagnosis == "FCD IIb", 2]` FCD IIb patients.


For these patients, we have only information on the diagnosis.  


## Controls - BrainSpan

```{r}
# Load metadata 

setwd(paste0(homedir, "Meu Drive/Doutorado/Deconvulucao/data/metadata_padronizada"))
meta_BS <- read_excel("std_meta_BrainSpan_FCD_Controls.xlsx") 
```



## Joining all of the metadata

```{r}
# bind and add info on the dataset
meta <- bind_rows(meta_disc %>% mutate(cohort = "Dataset 1"), 
                 meta_val %>% mutate(cohort = "Dataset 2"), 
                 meta_BS %>% mutate(cohort = "Control"))
```



# Signatures 

We'll look only at the signature CA first. 

```{r}
# Load data 

setwd("../CA")

ciber <- read.delim("CIBERSORTx_CA_FCD_all_Controls_BrainSpan_Rel.txt")


```



# Plotting it all together 

```{r}
# Join with metadata 
data <- ciber %>% inner_join(meta, by = c("Mixture" = "sample")) %>% 
  mutate(cohort =  factor(cohort, levels = c("Dataset 1", "Dataset 2", "Control")))


cell_types <- c("Neurons", "Excitatory", "Inhibitory", "Astrocytes", 
                "Oligodendrocytes", "Microglia", "Endothelia", "OPCs")

```

```{r, warning = F}

# We are using only the CA signature 


## Without quantile normalization 


melted_data <- reshape2::melt(data, measure.vars = cell_types)

plot <- ggplot(melted_data,
     aes(x = cohort, y=value,
         fill = diagnosis,
         color = diagnosis)) +
geom_violin(trim = FALSE, na.rm = TRUE,
            position = position_dodge(0.8), alpha = 0.7) +
  geom_boxplot(position=position_dodge(0.8),
             outlier.shape = NA, na.rm = TRUE, width = 0.25, color = "black") +
  stat_summary(fun = median, geom = 'crossbar', color = 'black', width = 0.2,
               position = position_dodge(0.8)) +
labs(x = "", y = "Proportion", fill = 'Subtype', color = 'Subtype') +
#stat_summary(fun.y=mean, geom="point", size=2, color="black", na.rm = TRUE) +
#scale_y_log10(labels=trans_format('log10', math_format(10^.x))) +
scale_fill_manual(values = c("lightblue", "purple", "orange") ) +
scale_color_manual(values = c("lightblue", "purple", "orange") ) +
facet_wrap(~ variable, ncol = 4, scales = "free_y") +
# expand limits so we can put asterisks for significance levels
scale_y_continuous(expand = c(0.2, 0)) +
 theme_bw() +
theme(axis.text.x = element_text(size=11),
      #axis.ticks.x = element_blank(),
      axis.text.y = element_text(size=12),
      axis.title.x = element_text(face="plain", colour="black", size=12),
      axis.title.y = element_text(face="plain", colour="black", size=12),
      #legend.title=element_blank(),
      panel.background=element_blank(),
      #panel.border=element_blank(),
      panel.grid.major=element_blank(),
      legend.text = element_text(face="plain", colour="black", size=12),
      legend.title = element_text(face="plain", colour="black", size=12),
      legend.position = "left",
      plot.title = element_text(face="plain", colour="black", size=14, hjust = 0),
      strip.text.x = element_text(face = 'plain', size = 10))

#print(plot)

#ggsave(filename = "plots_FewerControls_all/rel_CA_all_noQN.svg", plot = plot, width = 12, height = 6)


# Statistical tests

for (dat in c("Dataset 1", "Dataset 2")){
  
  data_test <- melted_data %>% filter(cohort %in% c(dat, "Control"))
  
  res <- lapply(unique(melted_data$variable), function(pop){
  
    # Kruskal-Wallis test 
    
    data_test_ct <- data_test %>% filter(variable %in% pop)
    
    kruskal.test(value ~ diagnosis, data = data_test_ct)
    
    # subset doesn't work here
    pwtest <- pairwise.wilcox.test(data_test_ct$value, data_test_ct$diagnosis, 
                         p.adjust.method = "none",
                         paired = F, 
                         exact = TRUE)
      
    table <- broom::tidy(pwtest)
    
    })
  
  names(res) <- unique(melted_data$variable)

  
  # Exporting 

#   lapply(names(res), function(pop){
# 
#      write.xlsx(as.data.frame(res[[pop]]),
#            file = paste0("p_values_FewerControls_all/rel_CA_all_noQN_pvalues_pairwise_wilcox.xlsx"),
#            row.names = F,
#            sheetName = paste(pop, str_remove(dat, "ataset "), sep = " - "),
#            append = T)
# 
# })

}


```

```{r}
# Save workspace 
save.image("FCD_AusFabiFilteredHoalnda_BrainSpan_Controls_all_together_clean.RData")
```

